---
title: "Lab 5"
author: "Renee LaManna"
date: "10/26/2021"
output: 
  html_document: 
    toc: yes
    theme: cerulean
    number_sections: yes
    code_folding: hide 
---

```{r setup, include = TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Confidence intervals

In this section, we'll learn to use the `t.test()` fxn as a way to return the confidence interval using the t-distribution 

## Example 1

Create a mock sample of obsv values, and find the confidence interval. Then we'll learn how to report that. 

```{r}
pika_mass <- c(82, 94, 110, 70, 86, 99, 102, 5000)

pika_mean <- mean(pika_mass)

pika_sd <- sd(pika_mass)

pika_ci <- t.test(pika_mass)
```

### tedious, risky, not great, not reproducable if changing a different number

Mean pika mass at Paiute Pass is 91.9 g $\pm$ 13.5 g 
(mean $\pm$ 1 standard deviation, n = 7) with a 95% CI of [79.4, 104.4] g.  

### A high intial investment, but safer & better way

In-line code referencing to stored objects

Use a single backtick on either side of a lowercase r to create an in-line bit of code.
If anything in my code changes, this kline will be updated with correct values
Rounded to 1 deciamal place

Mean pika mass at Paiute Pass is `r round(pika_mean,1)` g $\pm$ `r round(pika_sd, 1)` g (mean $\pm$ 1 standard deviation, n = `r length(pika_mass)` ) with a 95% confidence interval of [`r pika_ci$conf.int[1]`, `r pika_ci$conf.int[2]`] g.

# can call specific parts of t-test with (in console) pika_ci$conf.int or pika_ci$p.value

# Two-sample t-test

Use a 2-sample t-test to test null hypothesis that samples were drawn from pops with same mean (difference in means = 0)

- H0: Means difference = 0
- HA: Means difference is NOT = 0

This is a 2-sided t-test because no directionality is implied

## Example 1

```{r}
oak_trees <- c(29, 19, 22, 30, 35, 16, 20, 7)

pine_trees <- c(48, 32, 41, 40, 56, 70, 30, 19, 20)
```

Is there a signif difference in mean heights for oak trees and pine trees?
After we determine that t-test is the appropriate test based on assumptions, then we can run this test

```{r}
trees_t <- t.test(oak_trees, pine_trees)
```

type just trees_t into console to view results

This p-value of `r trees_t$p.value` means that there is `r (trees_t$p.value) * 100` % chance of finding sample means *at least as different as those I found* by random chance if they were drawn from populations with same mean height

My means difference is -2.7 away from 0.

We would reject the null hypothesis because there is a 1.8% chance that....


## Example 2

Using the `mpg` dataset to compare city gas mileage for SUVs and compact cars.
Many different ways to do this....

Make subset called compact_suv that only contains obsv wehre class is compact OR SUVS
- unique(mpg$class) in console to view all the different vectors in a single column unique(dataframe$column)

```{r}
compact_suv <- mpg %>% 
  filter(class %in% c("suv", "compact")) 
# check and make sure works in console/easier than searching through datatset   unique(compact_suv$class)
```

# want to see suv and compact separately, here we use facet_wrap() to split up in each class
# can adjust bin, default = 30
# histogram only needs x value

# next create quantile quanitle plot
- perfectly normal distributions would have linear relationshiops
- should only be used in combo with histograms

```{r}
ggplot(data = compact_suv, aes(x = cty)) +
  geom_histogram(bins = 12) +
  facet_wrap(~ class)

ggplot(data = compact_suv, aes(sample = cty)) +
  geom_qq() +
  facet_wrap(~ class)
```
# by group_by() that exists within class

```{r}
car_stats <- compact_suv %>% 
  group_by(class) %>% 
  summarize(mean_city = round(mean(cty),1),
            sd_city = round(sd(cty), 1),
            sample_size = n())
car_stats
```

```{r}
compact_sample <- compact_suv %>% 
  filter(class == "compact") %>% 
  pull(cty)

suv_sample <- compact_suv %>% 
  filter(class == "suv") %>% 
  pull(cty)

cars_t <- t.test(compact_sample, suv_sample)
```

Reject null hypothesis of equal mean city gas mileages: stats speak
in paper: mean gas mileage (say means) for compact cars and SUVs differ significantly (t(`r cars_t$parameter`) = `r cars_t$statistic`, p < 0.001, $\alpha$ = 0.05).
